### 2011-08-23 response/revision_plan.tex
comments regarding parameter sensitivity

\documentclass{article}
\usepackage{dcj}
\usepackage{ulem}

\begin{document}

\dcjtitle{Seqbias Paper}{Plan for Revision}{Daniel Jones}

\begin{enumerate}

\item \sout{Demonstrated the results are not particularity sensitive to the
complexity penalty, drawing a plot similar to the one is Supplementary Section
3, but varying the penalty rather than the number or reads.}

\item \sout{Point out that the fact that the method is effective over diverse sets of
data argues that the results are not sensitive to any of the parameters used.}

\item Note that the Hansen's brute-force method was trained using many more than
600,000 reads on some of the data sets.

%\item (Maybe) Draw a curve showing how the performance of Hansen's brute-force
%method changes as a function of the number of reads fed into it.

% This is sort of implied by the FRT-seq data
\item Make very clear that in there is no bias present in the data, a sparse or
empty model will be trained.

\item \sout{Include a figure from the genome browser showing an example uniformity of
coverage being increased over a gene or exon.}

\item Use the MAQC set to demonstrate the following:
\begin{enumerate}
\item Increased agreement between bias corrected sequencing data and qPCR.
%\item Increased agreement between replicates. % Won't work here
\end{enumerate}

\item \sout{Examine the bias in new methods without a PCR-amplification step.}

\item \sout{Typos mentioned by reviewr 1.}

\end{enumerate}

\end{document}