### 2011-09-29 response/response.tex
more revisions

\documentclass{article}
\usepackage{amsmath}
\usepackage{dcj}

\begin{document}

We appreciate the time and effort extended by the anonymous reviewers and the
editorial staff to evaluate the original version of this manuscript, which was
submitted on July 8. Upon review, the second of the two reviewers raised some
concerns leading to its rejection. However, we believe that the bulk of these
concerns are either incongruous with the content of the manuscript or
addressable with very little effort, and therefore do not warrant its rejection.
Here we have provided a detailed response to these comments, and have included
additional material in an effort to clarify these points.

The second reviewer raised three major concerns. We summarize our
responses to these issues here.  The first concern raised is
that the method we present is unjustifiably complex, involving a ``plethora of
tuning parameters'', and that the much simpler method of Hansen, et al., would
perform as well if trained on enough data.  In the data we presented, the simple
method developed by Hansen, et al., performed poorly regardless of the number of
reads used in training, which varied between hundreds of thousands to tens of
millions.  The methods presented by Li, et al., are more effective, yet involve
more parameters, and require more time to train, so that our method, while often
being more effective, also offers a significant simplification.
Additionally, we believe the structure of our models give some
theoretical insight into these differences in efficacy.

Secondly, it is asserted that we did not demonstrate that uniformity of read
coverage is increased after correcting for bias using our method.  Yet, in
section 3.2 we had presented a rigorous statistical analysis over a large and
diverse collection of data sets showing that this is indeed the case. This
section of the paper was perhaps overlooked by the reviewer, as it received
no mention in this criticism.

Finally, the reviewer doubts the biological relevance of the bias correction
methods. Having demonstrated an increase in uniformity (section 3.2) and a
decrease in nucleotide frequency divergence (section 3.1), we believe a
compelling case was made that correcting for bias increases the accuracy of
quantification. In this revision we have supplied another line of evidence:
analysis showing that correlation between RNA-Seq and qRT-PCR is increased after
correcting for bias.

Under these circumstances we respectively request that the manuscript, in this
revised form, be reconsidered for publication.  Detailed responses to
the reviewers concerns are presented below.

\subsection*{Response to Reviewer 1}

\begin{quote}
(1) Page 3, 2.2.1 mentions a correction for bias introduced by PCR amplification.
RNA sequencing methods are available where PCR amplification is not used.
Correcting for bias in these cases removes true duplication.  Could the authors
please comment on this?
\end{quote}

First, we emphasize that PCR bias is only one of a number of technical steps
involved in sequencing that potentially introduce bias.  Our method is agnostic
to the source of bias, and because model complexity is penalized, a very sparse,
or even empty model will be trained on a data set without bias.  Thus, if PCR
amplification is responsible for a majority of the bias, and an RNA-Seq method
without PCR amplification is used producing largely bias-free data, then using
the method to correct for bias will have little if any effect.

As a concrete example, we have included new data examining the FRT-Seq protocol
of Mamanova, et al., in which no PCR-amplification is performed during library
preparation. We observe that bias is still present in this data set, albeit at
lower levels than other uncorrected RNAseq data examined.  Our method accounts
for this reduced bias by constructing a much sparser model, relative to the
other RNA-Seq data sets we have examined.

To further explore the point raised, we repeatedly trained our model on sets of
10,000 random simulated reads, drawn uniformly from within exons. We repeated
this 100 times, and in each trial an empty model was trained.  Reducing the
number of reads increases the risk of a spuriously non-empty model, but
repeating this evaluating with only 100 reads led to only 3 out of 100 models
being non-empty. This is indicative that the default BIC scoring, as predicted
by statistical theory, is adequately protecting against false discovery of
``bias'' when there is none.

\subsection*{Response to Reviewer 2}

\begin{quote}
The ms presents an approach to estimating the effect of local sequence
context on the rate of seeing aligned reads at each position within a
transcript in RNA-Seq data. The method is inspired by Hansen et al.'s,
which simply determines this bias by comparing heptamer frequencies.

Compared to the brute force approach of Hansen, who estimated model
parameters (the density of a joint distribution) directly from
heptamer frequencies, here a Bayesian network approach is taken, which
allows for a sparser parameterisation. However, the cost for this are
a plethora of tuning parameters (e.g. the complexity penalty, data
sampling, $p_{max}$, $d_{max}$) and a greedy algorithm with apparently no
control of the quality of its solution.
\end{quote}

The method has precisely the following six parameters,
\begin{enumerate}
\item The number of reads the method is trained on.
\item The size of the window surrounding read starts over which the model should
be trained.
\item The maximum in-degree (i.e., $p_{\text{max}}$)
\item The maximum edge distance (i.e., $d_{\text{max}}$).
\item The standard deviation at which background sequences are sampled.
\item The model complexity penalty.
\end{enumerate}

Parameters (1), (2), (3), and (4) are introduced only to reduce the CPU time
needed to train the algorithm. Each has the same very simple interpretation: a
larger value leads to a potentially better solution, at the cost of increased
training time.

We mention these parameters in the manuscript only for the sake of transparency.
They generally require no intervention or tuning, being essentially an
implementation detail.

Parameters (5) and (6) are more important to the quality of the solution, yet we
have demonstrated, with no special tuning, the method is effective across a
diverse collection of data sets. In this revision we have included new data
(Supplementary Section 2) showing explicitly that the quality of the solution is
remarkably insensitive to values of these two parameters over a range spanning 4
orders of magnitude.

It should also be noted that, while the brute-force method of Hansen, et. al.,
has fewer tuning parameters, the other methods evaluated in the manuscript (the
GLM and MART methods of Li, et. al.), involve more parameters than the method we
have described. (For example, training the MART method involves parameters
controlling the complexity penalty, the number of trees, the lower bound
threshold for read counts, in addition to others that are not described or
easily changed from the values divined by the authors.  Additionally, their
method requires selection of a set of annotated genes for training.)  Our method
therefore presents a simplification over competing methods.

\begin{quote}
To justify this cost, the
authors make the argument that their approach avoids the potential
overfitting of a brute force approach such as Hansen's. However, I am
not convinced: they evaluate both methods, theirs and Hansen's, on
'more than 600,000 reads', yet this seems unrealistic: current
datasets have hundreds of millions of reads, and that may just solve
the estimation precision problem in the simple brute force approach.

(I appreciate that the implementation of the brute force approach in
the Genominator package is ridiculously inefficient, but that can
easily be fixed by more competent programming.)
\end{quote}

We regret that we were not more explicit on this point. At least
600,000 reads on each data set, but as many as 23 million were used:
see the revised table X. (As noted in the paper, we did reimplement
the Genominator method in order to do this.)

We do not wish to ``pick on'' Hansen, et al.  Theirs was a pioneering
paper and presents a simple, intuitive approach for tackling the
problem they identified.  Nevertheless, their brute-force method
performed poorly on every data set we tested, even the largest ones,
suggesting that simply throwing more data at a suboptimal solution
will not adequately solve the problem.  In our opinion, the results we
presented were unambiguous: the other methods perform far better with
far fewer reads.

%Moreover, even if the brute-force method were effective if trained
%with many hundreds of millions of reads, many (possibly most) existing
%data sets are not this large, rendering it generally impractical.

Understanding this failure of a simple, appealing method is important
for making progress, and we feel that our paper provides a theoretical
basis for understanding these differences in efficacy, and we have
amplified these points in the revision.  Intuitively, one expects
model accuracy to increase as the number $n$ of training reads
increases, but to decrease as the number $r$ of model parameters
grows, since the training data is ``spread more thinly.''  For the
particular case of estimating a multinomial distribution (as in
Hansen), Supplementary Section YY quantifies this intuition very
precisely, showing that the expected KL divergence of the learned
model with respect to the true distribution converges to $(r-1)/(2n)$,
where $r=4^7=16384$ is the number of model parameters learned.  The
same analysis does not directly apply to the MART, GLM or BN models,
but we expect their behavior to be qualitatively similar.  Thus, to
the extent that a more parsimonious model such as ours adequately
captures the structure of the data, it is potentially orders of
magnitude more accurate for a given quantity of training data, simply
by virtue of having only a few hundred parameters learned from counts
(Figure 3), as opposed to tens of thousands.  [KL divergence between
two distributions is inversely proportional to the number of samples
one needs to confidently test the hypothesis that the samples arise
from one distribution versus the other; see Supplementary Section XX.]
%% TODO: why the brackets here?

Of course, training data is plentiful in NGS data sets, so,
pragmatically, the above analysis probably identifies a contributor
but not the root cause of Hansen et al.'s relatively poor performance
in our experiments.  Rather, we believe the problem is that bias is
not confined to the initial heptamer.  The model structures
illustrated in Figure 3 show that nucleotides well outside that region
are often biased.  (The randomization study sketched above shows that
our model building algorithm is not prone to spurious discovery of
overly-complex models.)  Furthermore, these examples show that the
read sampling is often biased by positions \emph{in front of the
  read}, which are not currently considered by Hansen.  Their method
could be modified easily to consider $k$-mers spanning the read start,
but the entropy analysis shows that extending it to $k \gg 7$ is
infeasible, since the exponential growth in $r=4^k$ will destroy the
model's accuracy even in large data sets.  E.g., $k \ge 20$ would be
necessary to cover the biased regions in all data sets shown in Figure
3, necessitating trillions of training reads to build an adequate
model by Hansen's method.

\begin{quote}
The major weakness of the ms, in my view, is the choice of 'figures of
merit' that the authors made (Fig.1, Sections 3.1, 3.2). What I would
have expected to see are

1.) Plots closer to the data such as Supplementary Fig.5 of
reference [1]. That figure shows highly uneven coverage of reads
within exons.
- Does this unevenness go away with the proposed adjustment?
\end{quote}

During preparation we chose not to include a figure such as the one suggested
by the reviewer as we believed it would not be legitimately informative. It
would be far too easy to cherry-pick a miniscule genomic interval that makes
even a very shoddy method appear favorable.

Instead, we presented data (in Section 3.2) that answers the
reviewer's question (i.e., ``does this unevenness go away with the
proposed adjustment?'') in rigorous and comprehensive manner. We
implemented a Poisson regression metric, which quantitatively measures
``uniformity'' by calculating the likelihood that the reads sampled
across a given exon were sampled from a uniform distribution.  Using
this, we performed a large cross-validation study proving the
effectiveness of our method at increasing uniformity across thousands
of exons drawn from a diverse group of data sets, rather than
qualitatively across one exon.  In broad terms, our method increases
likelihood of the uniform model by XXX-fold across thousands of exons
(Figure 6).

In this revision, as suggested, we have revised Figure 2 to show an
example of this increase in uniformity.  We believe the figure is a
nice, graphic illustration of the benefits of our method and thank the
reviewer for causing us to reconsider it.  We stress, however, that
this figure represents a minute subset of the analysis that had
previously been presented, and is included only as an illustration.
The results in sections 3.1 and especially in 3.2 are far more
comprehensive.

\begin{quote}
- Also, as is evident from that figure, for some of the experimental
protocols, some regions within highly expressed exons get vanishingly
small read coverage.  These are the really hard situations, and much
more worrying than instances where coverage just fluctuates a bit.
How does the proposed scaling adjustment of Section 2.1 work in these
places (i.e. when $s_{i}$ is very small)?
\end{quote}

%% TODO

%The revised Figure 2 also graphically illustrates the range of
%adjustment factors, which genome-wide are very cool and we hope we can say something soothing about.......

\begin{quote}
2.) A 'bottom-line', biologically relevant figure of merit: E.g., does
the identification of differentially expressed genes improve? The
clustering of gene expression profiles? The identification of
transcript isoforms?
\end{quote}

In Section 3, we demonstrate, after correction, a significant improvement in
the two obvious anomalies associated with the issue of sequence bias:
non-uniform read coverage, and divergent nucleotide frequencies. We assess these
aspects in a statistically rigorous manner and with a large sample size. We
believe that this analysis presents a compelling case that bias correction is
worthwhile, and that the method we present is competitive with, and quite often
superior to existing methods.

In this revision, we have included more data, showing that in addition to the
improvements in uniformity and nucleotide frequencies, increased correlation
between RNA-Seq and qRT-PCR, after adjusting for bias in RNA-Seq. These results
are consistent with, and reinforce our previous results.

\begin{quote}
Minor points:
------------
1. The ms seems confused between classification and regression.
Reference [2] might be helpful to resolve some of that.
\end{quote}

The term ``classification'' was used in the context of noting that the algorithm
is adapted from one typically used in classification. It is correct to say that
the method here performs regression, and we have tried to clarify this language
in this revision.

\begin{quote}
2. The chosen Bayesian network is rather complicated, and a spectrum
of alternative methods exists between that and the brute force
approach of Hansen (say, lasso or svm regression). In absence of a
more meaningful benchmark, the choice of method seems arbitrary.
\end{quote}

Our motivation for adopting a Bayesian network model is the very
natural probabilistic interpretation that we have described in Section
2.1. We disagree with the reviewer's opinion that Bayesian networks
are inherently more complicated than Lasso regression or support
vector machines.  In this application, the latter two methods would
generate output that is arguably less transparent and interpretable,
and would neither be significantly simpler to implement, nor involve
fewer parameters.  Our hypotheses (above) concerning the reasons for
the weakness of Hansen's method illustrate the benefit of this
approach; we do not believe svm regression, say, would have yielded
similar insights.

Naturally, it is possible that another model might perform as well or better
than the one presented here, yet it is unrealistic or impossible to prove that
this is not case. What we have demonstrated is a model that is well reasoned,
efficiently implemented, and effective at reducing sequence bias in RNA-Seq.

%\subsection*{Other Changes}

%We have modified the title to more clearly acknowledge the existence of
%the prior work upon which we build.

%...more...?

\subsection*{Conclusion}
In summary, we believe the revised manuscript addresses all concerns
raised by the reviewers, and provides a substantive improvement to the
state of the art for an important problem.  We respectfully request
that the manuscript, in this revised form, be reconsidered for
publication.

\end{document}